---
title: "Water_Fowl_Testing_Area"
author: "Samantha Hunter"
date: "11/19/2021"
output: html_document
---


JSTOR LINK TO STUDY:https://www-jstor-org.prox.lib.ncsu.edu/stable/3783692?seq=2#metadata_info_tab_contents

The data set is from US Fish and Wildlife services. I took the most recent six years of data that was available (2010-2016). Information about the data set can be found (here)[https://migbirdapps.fws.gov/mbdc/databases/afbws/aboutafbws.htm]

The survey's primary target is to estimate the breeding population size of mallards, black ducks, wood ducks, and Canada geese so that is the only data I will be looking at. 
```{r}
#install.packages("tidyverse")
library(tidyverse)

# use to fix the date values from character to actual date
#install.packages("lubridate")
library(lubridate)

#install.packages("caret")
library(caret)


fowls <- read.csv("duck_data.csv")

fowls$STRATUM = as.character(fowls$STRATUM)

birds <- fowls %>% 
  mutate( 
        # Combining the old goose counts with the new goose counts, 
        # if the new count exists, then that is what we will use, otherwise we'll use the old count...
        # not the best practice but good enough for now
        CanadianGoose = if_else(!is.na(CAGO_TIBN), CAGO_TIBN, CAGO_TIB)) %>%
  
  rename(Year = YEAR, State = STATE, TimeOfDay = CKTYPE, Plot = PLOT, Date = DATE,
          Stratum = STRATUM, WetHab = WETHAB, Handfeed = HANDFEED, 
          Mallard = MALL_TIB, 
          AmBlackDuck = ABDU_TIB, 
          WoodDuck = WODU_TIB) %>%
  
  # Removing extra columns - Have the TIBN and TIB data else where
  # The pairs of birds seems to have been calculated based on the number of birds seen
  # so I am just using the birds variables -> Maybe come back to automate for Pairs
  select(!CAGO_TIB, -CAGO_TIBN, -ends_with("_TIP")) %>%
  
  # Pivoting the table so each observation denotes the count of one species
  # (making it easier to filter)
  pivot_longer(cols = 9:12, names_to = "Species", values_to = "Count") %>%
  
  # Only keeping explicit zeroes. The data was square so I assume those were introduced
  # by the structure as there are plenty of zeroes as well
  filter(!is.na(Count))


# Fixing the dates to actual date instead of character strings
birds$Date <- mdy(birds$Date)

# This has all years
# unique(birds$Year)

# The overwhelming majority of these are zeroes (105526/134444, this is about eighty percent of the data is zeroes)
# Removing these zeroes removes all observations for 2010 and 2012. 
count_zeros <- birds %>% filter(Count == 0)

# These are just records where birds were seen
bird_count <- birds %>% filter(Count != 0)
# No 2010 or 2012, meaning that the only recorded counts those years is 0 for all species
# that's not very helpful anyways
# unique(bird_count$Year)

# # Just looking at these dates that are missing or a null value, 
# # they seem to be all over the place (lots of states, years, ect)
# # so I'm just going to make a note of that... For now I'm just not going to 
# # use this variable
# missing_date <- birds %>% filter(is.na(Date)|Date == "1900-01-01")
# # NOTE: There are 30310 records who don't have the date recorded and it looks like they are all cohorts
# 
# # Just for the records we look at, only 3891/28918 are missing the date
# missing_date_count <- bird_count %>% filter(is.na(Date)|Date == "1900-01-01")
# # so I'm just going to get rid of that 13% of data as well
bird_count <- bird_count %>% filter(-is.na(Date)|Date != "1900-01-01")

# Finally, I can see that I have two records listed as 20/04/1930 in Date, while Year reads 2003
# so those dates will become 20/04/2003
bird_count[bird_count == "1930-04-20"] <- as.Date("2003-04-20")

# Defining bird sightings by season for a bar plot
bird_season <- bird_count %>% mutate(
  season = cut(yday(bird_count$Date - days(79)), 
            breaks = c(0, 93, 187, 276, Inf), 
            labels = c("Spring", "Summer", "Autumn", "Winter"))
)
# Birds are overwhelmingly seen in spring, not very useful data
table(bird_season$season)

# What about months?
# Nearly all birds were sighted in April and May (breeding season?)
hist(month(bird_count$Date))

# about half and half
table(bird_count$WetHab)

# Overwhelmingly skewed towards no handfeeding around
table(bird_count$Handfeed)

# Overwhelmingly skewed towards 1
table(bird_count$TimeOfDay)

# Finely distributed 
table(Stratum)

# cut divides the range of x into intervals and codes the values in x according to which interval they fall
# yday and days are both from the lubridate package, yday returns the day of the year


# Here we are showing that we should remove the NA counts. There are records with 0 in line with NA
# There are 287 of these so it is a really small portion of the data records (per site)
# missing_data <- birds %>% filter(is.na(AmBlackDuck)|is.na(Mallard)|is.na(WoodDuck)|is.na(CanadianGoose))
```


Variable meaning [https://migbirdapps.fws.gov/mbdc/databases/afbws/aboutafbws.htm]

Important Info: All plots are checked annually in 1993

Year: Survey Year
State: State in which sample plot is contained
Stratum: Physiography stratum in which plot is contained- please see map on JSTOR, the numbers indicate what type of environment the area is (eg scrub-shrub estuarine wetlands, estuarine tidal flats, ect)
Plot: ID Variables if you sort data by Year-State-Stratum-Plot

### What I'm Interested In/Questions
1. General Trend of Bird Population
2. Can you estimate next year from previous years? (split data into 2010-2014, predict for 2015)?


# Data Exploration

### Tables
```{r}
# Just exploring the distribution of counts for each bird
# birds$CountCat = if_else(birds$Count == 0, "None", 
#                     if_else(birds$Count == 1, "Solo", 
#                       if_else(birds$Count == 2, "Pair", 
#                         if_else(birds$Count >= 3| Count < 30, "Gaggle", "Flock"))))
?subset

summary_table <- bird_count %>% group_by(Species,Year) %>%
  summarize(Mean = mean(Count), Median = median(Count), Range = paste0(min(Count), "-", max(Count), IQR = IQR(Count)))
summary_table

# Have this be rendered in the data exploration tab
bird_count %>% group_by(Species, Year) %>%
  summarize(Mean = mean(Count), Median = median(Count), Range = paste0(min(Count), "-", max(Count)))


unique(birds$Stratum)
# 12
# 27
# 99 - Saltmarsh Stratum

```


### Graphs n Such

###### Need at least one graph for each variable
Count is response variable
- Year = year on x variable, count on y variable
- State
- Stratum
- Date
- Time of Day = histogram?
- WetHab - histogram
- Handfeed 
- Species = same as year?

```{r}
# Heat map of date, first remove the missing dates
# Trendline of Time vs Count, each separate line available for species, scatterplot dots colored by species
ggplot(bird_count, aes(x = Year, y = Count, colour = Species)) +
  geom_jitter() + 
  stat_smooth(data = subset(bird_count, Species == "AmBlackDuck", 
                            method = "lm", se = T)) +
  stat_smooth(data = subset(bird_count, Species == "CanadianGoose", 
                            method = "lm", se = T)) + 
  stat_smooth(data = subset(bird_count, Species == "Mallard", 
                            method = "lm", se = T)) + 
  stat_smooth(data = subset(bird_count, Species == "WoodDuck", 
                            method = "lm", se = T))

ggplot(bird_count) +
  geom_bar(aes())

subset(bird_count, 7 <= Count & Count <= 400 )

ggplot(data = subset(bird_count, 0 <= Count & Count <= 400 )) +
  geom_jitter(aes(x = Year, y = Count, shape = Species, color = State))

```


# Modeling
Don't use plot or date in the linear modeling. Date has too many missing values (about a third of the data is missing date) for me to feel comfortable using it as a predictor. I also am choosing not to use plot as it seems to be an identifying variable instead of a descriptive variable. 

### Splitting the data/Preprocessing
```{r}
# Making a split so we get a representative number of each species in the same
# distribution as they appear in the data
# set.seed(15)
# 
# mallard <- birds %>% filter(Species == "Mallard") %>% select(!Species)
# 
# train <- sample(1:nrow(mallard), size = nrow(mallard)*0.7)
# test <- setdiff(1:nrow(mallard), train)
# 
# mallardTrain <- mallard[train, ]
# mallardTest <- mallard[test, ]

```


Use if we decide to use all birds species at once, but I think I only want to let you predict if you
choose a species...
```{r}
trainIndex <- createDataPartition(bird_count$Species, p = 0.7, list=FALSE)

birds_train <- bird_count[trainIndex,]
birds_test <- bird_count[-trainIndex,]

preProc <- preProcess(birds_train, method = c("center", "scale"))
birds_std_train <- predict(preProc, birds_train)
birds_std_test <- predict(preProc, birds_test)
```



### Linear Model
```{r}
fit <- lm(Count ~ Year + State + Stratum + TimeOfDay + WetHab + Handfeed + Species, data = birds_train)
summary(fit)
plot(fit)

lmFit1 <- train(Count ~ Year + State + Stratum + TimeOfDay + WetHab + Handfeed + Species,
                data =birds_train, method = "lm", trControl = trainControl(method = "cv", number = 5), 
                na.action = na.exclude)

# Model Comparison

summs[["adj.r.squared"]]

summs <- summary(lmFit1)




```




### Tree
```{r}
rf_SigPred <- train(Count ~ Year + State + Handfeed + Species,
             data = birds_train, method = "rf", 
             trControl = trainControl(method = "cv", number = 5),
             tuneGrid = data.frame(mtry = c(1, 2, 3, 4)))
rf_SigPred
plot(rf_SigPred)


rfFit <- randomForest(Count ~ Year + State + Handfeed + Species,
                      data = birds_train, mtry = 1:4, 
                      ntree = 200, importance = TRUE)
rfFit
tree(rfFit)
# Plot representing the 
plot(rfFit, xlab = "Number of Predictors in RF Model", ylab = "Root Mean Square Error (Found through CV)", 
     main = "Minimizing RMSE")
```



### Random Forest 
```{r}

```








